---
title: "Rental Pricing in KL"
author: "Ken Law"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Rental Prices in Kuala Lumpur (02/12/2023)
  
## Description
This dataset is provided by ARIEWIJAYA on Kaggle. It is scraped from local rental advertising listing websites in Kuala Lumpur. It contains a few listings from the surrounding outskirt areas of KL. There are 9991 listings total with 13 variables. Primary key is set as 'ads_id' by the dataset owner.
  
## Variables
1. Advertising ID: Unique ID assigned to each *primary key*
2. Property name: (Character)
3. Completion year: (Date/Integer)
4. Monthly rent: Currency in Ringgit Malaysia  (Currency/Numeric)
5. Location: 'City - Region' (Location/Character)
6. Property Type: Categorical e.g. Apartment, Condo (Factor)
7. Rooms: (Integer)
8. Parking: (Integer)
9. Bathroom: (Integer)
10. Size: Calculated in sq. ft. (Integer)
11. Furnished: Fully, Partially, or Not Furnished (Factor)
12. Facilities: Listed amenities and facilities provided (Factor)
13. Additional Facilities: Air-con, Public Transport, etc (Factor)
  
  
## Business Task

**Market Analysis:**  
Analyze the location, property type, size, and completion year columns to understand the current market trends and demand for specific types of properties in certain areas.

**Pricing Optimization:**  
Use the completion year, location, property type, size, furnished, and facilities columns to determine the optimal price for a property.

**Customer Segmentation:**   
Analyze the facilities and additional facilities columns to identify patterns in customer preferences and segment the market accordingly.

**Property Selection:**   
Use the completion year, location, property type, size, furnished, facilities, parking, bathroom, and rooms columns to identify properties that meet specific customer needs and preferences.

**Advertising and Marketing:**   
Use the advertising ID, property name, completion year, location, property type, size, furnished, and facilities columns to create targeted advertising campaigns and marketing strategies.

**Inventory Management:**   
Use the furnished column to track which properties are fully, partially, or not furnished and manage inventory accordingly.

**Maintenance and Improvement:**   
Use the completion year and facilities columns to identify areas where improvements or maintenance may be needed to keep the property competitive in the market.

**Revenue Forecasting:**   
Use the monthly rent and size columns to forecast potential revenue and evaluate the profitability of a rental property.

**Performance Tracking:**   
Use the monthly rent, size, parking, bathroom, and rooms columns to track the performance of a property over time and identify areas for improvement.

**Risk Assessment:**   
Use the location, completion year, and property type columns to assess the risk of investing in a specific property.

---
  
## Data Processing 
Loading the packages:
```{r}
library(here)
library(tidyverse)
library(dplyr)
library(readr)
library(stringr)
library(janitor)
library(skimr)
```

Open CSV file:
```{r}
mudah_apartment_clean <- read_csv("C:/Users/lawke/Desktop/portfolio_2023/rental-pricing-in-kuala-lumpur-malaysia/mudah-apartment-clean.csv")
```

Quick look:
```{r}
skim_without_charts(mudah_apartment_clean)
glimpse(mudah_apartment_clean)
head(mudah_apartment_clean)
str(mudah_apartment_clean)
```
  

Split location into city and region:
```{r}
mudah_apartment_clean <- mudah_apartment_clean %>%
  mutate(city = str_split(location, "-", simplify = TRUE)[, 1],
         region = str_split(location, "-", simplify = TRUE)[, 2]) %>% select(-location)  
```

## Data Validation
```{r}
class(mudah_apartment_clean$prop_name)
class(mudah_apartment_clean$completion_year)
class(mudah_apartment_clean$monthly_rent)
class(mudah_apartment_clean$city)
class(mudah_apartment_clean$region)
class(mudah_apartment_clean$rooms)
class(mudah_apartment_clean$parking)
class(mudah_apartment_clean$bathroom)
class(mudah_apartment_clean$size)
class(mudah_apartment_clean$furnished)
class(mudah_apartment_clean$facilities)
class(mudah_apartment_clean$additional_facilities)
```
  
  
## Data Cleaning and Transformation
Convert monthly_rent (character to numeric):
```{r}
mudah_apartment_clean <- mudah_apartment_clean %>% 
  mutate(monthly_rent = as.integer(str_replace_all(mudah_apartment_clean$monthly_rent, c("RM" = "", "per" = "", "month" = "", " " = ""))))


class(mudah_apartment_clean$monthly_rent)

```


Convert size (character to integer):
```{r}
mudah_apartment_clean <- mudah_apartment_clean %>% 
  mutate(size = as.integer((str_remove(size, "sq.ft."))))
class(mudah_apartment_clean$size)
```

Convert furnished (character to factor): 
```{r}
mudah_apartment_clean <- mudah_apartment_clean %>% 
  mutate(furnished = as.factor(furnished))
class(mudah_apartment_clean$furnished)
```

Remove Properties with no names (NA):
```{r}
mudah_apartment_clean <- subset(mudah_apartment_clean, !is.na(mudah_apartment_clean$prop_name))
```

---

# Analysis

Which regions in KL are the most expensive?
```{r}
expensive_region <-  subset(mudah_apartment_clean, monthly_rent > 200000)
ggplot(data = expensive_region, aes(x = expensive_region$region , y = expensive_region$monthly_rent)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Region", y = "Monthly Rent")
```

Which rental properties have the highest rent?
```{r}
mudah_apartment_clean <- mudah_apartment_clean %>%  arrange(desc(monthly_rent))
head(mudah_apartment_clean)
```
> 1. Sunway Vivaldi  = RM 2,400,000
> 2. The Tamarind = RM 580,000
> 3. United Point Residence @ North Kiara = RM 580,000
> 4. Parkhill Residence Bukit Jalil = RM 580,000
> 5. The Haute = RM 550,000
> * Monthly rent
  
Which rental properties have the lowest rent?
```{r}
tail(mudah_apartment_clean)
```
> 1. Setapak Ria Condo = RM 80
> 2. Idaman Sutera = RM 90
> 3. Platinum Lake PV 15 = RM 100
> 4. Platinum Hill PV 2 = RM 100
> 5. Platinum Lake PV 8 = RM 100
> * Monthly rent